---
title: "Assignment 4 – Webscraping 1"
format: html
editor: visual
---

## 1. Scraping Wikipedia Foreign Reserve Data

In this assignment I use **rvest** to scrape the table of foreign-exchange reserves from Wikipedia and then clean the resulting data frame.

```{r}
library(tidyverse)
library(rvest)
library(stringr)
library(lubridate)

url <- "https://en.wikipedia.org/wiki/List_of_countries_by_foreign-exchange_reserves"

wikiforreserve <- read_html(url)
```

## 1.1 Extract the first table (country-level reserves)

```{r}
# get all tables from the main content area
all_tables <- wikiforreserve |>
  html_nodes(xpath = '//*[@id="mw-content-text"]/div/table')

# first table = country-level reserves
foreignreserve <- all_tables[[1]] |>
  html_table(fill = TRUE)

# keep only the first 6 columns (the main country table)
fores <- foreignreserve[, 1:6]

# give those 6 columns clear names
names(fores) <- c("Rank", "Country", "Forexres", "Date", "Change", "Sources")

head(fores)
```

## 2. Cleaning the Data Frame

### 2.1 Clean the Date variable

```{r}
fores <- fores |>
  mutate(
    Date_clean_chr = str_split_fixed(Date, "\\[", n = 2)[, 1] |> str_trim(),
    Date_clean = dmy(Date_clean_chr)
  )

fores |>
  select(Country, Date, Date_clean_chr, Date_clean) |>
  head()
```

### 2.2 Remove unneeded rows and clean numeric fields

```{r}
fores_clean <- fores |>
  mutate(
    Forexres_usd = parse_number(Forexres),
    Change_num   = parse_number(Change)
  ) |>
  select(
    Country,
    Forexres_usd,
    Date = Date_clean,
    Change = Change_num
  )

head(fores_clean, 10)
```

## 3. Modifying the Program to Scrape Other Tables

```{r}
all_tables_df <- wikiforreserve |>
  html_table(fill = TRUE)

length(all_tables_df)
```

```{r}
table2 <- all_tables_df[[2]]
table3 <- all_tables_df[[3]]

head(table2)
head(table3)
```

```{r}
get_reserve_table <- function(idx) {
  tab <- all_tables_df[[idx]]
  tab
}

example_table <- get_reserve_table(2)
head(example_table)
```

## 4. Data Plan for Acquiring Web Data for Research

### 4.1 Define the research question
Identify variables, scope, and intended analysis.

### 4.2 Identify data sources  
Prefer official sources (IMF, World Bank, OECD).

### 4.3 Choose collection method  
rvest for tables, httr/jsonlite for APIs.

### 4.4 Cleaning workflow  
Standardize names, convert numeric/date fields, remove footnotes.

### 4.5 Storage & documentation  
Use data/raw and data/clean, maintain a data dictionary.

### 4.6 Update strategy  
Re-scrape periodically and check layout changes.

### 4.7 Ethics  
Respect robots.txt and cite data sources properly.

## 5. Summary

This assignment involved scraping with rvest, cleaning variables, extracting multiple tables, and designing a research data acquisition plan.
